---
title: 'Washington Data Codebook'
output: 
  html_document:
    toc: true # table of content true
    toc_float: true
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: readable  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE)
```

```{r load data, results='hide'}
les_packages <- lapply(c('prophet','data.table','tidyverse','ggplot','ggthemes','DT','readxl',
                         'reshape2','fst'),require, character.only = TRUE)
## load files
input_path = '../data/raw/'
output_path =  '../data/processed/'

#utils
process_col = function(x){
  y = gsub('[^[:alnum:] ]', '', x)
  y = gsub('\\s+',' ', y)
  y = tolower(gsub(' ','_',y))
  return(y)
}

check_sign = function(x){
  if(x==0|is.na(x)){
    return(0)
  } else if(x > 0){
    return(1)
  } else {
    return(-1)
  }
}

gen_growth = function(x,x_lag){
  #only handle numbers with same sign
  if(check_sign(x)!=check_sign(x_lag) | check_sign(x_lag)==0) return(check_sign(x) - check_sign(x_lag))

  y = (x-x_lag) / abs(x_lag)
  return(y)
}

median_interval = function(x,p=0.05){
  sort(x)[qbinom(c(p/2,1-p/2), length(x), 0.5)]
}

APE = function(actual,pred){
  if(is.nan(actual) | is.na(actual)) return(NA)
  if(actual!=0){
    return(abs(pred-actual) / abs(actual))
  } else {
    return(NA)
  }
}

log_trans = function(x){
  if(x>0){
    return(log(x))
  } else{
    return(NA)
  }
}
```

# Executive Summary

This codebook demonstrates the data ingestion and processing from Washington Department of Health's [Summary Financial Data](https://www.doh.wa.gov/DataandStatisticalReports/HealthcareinWashington/HospitalandPatientData/HospitalFinancialData/SummaryFinancialData). This includes:

* Balance Sheet
* Income Statement
* Financial Ratios
* FTE
* Rate Analysis
* Volume Trends

# Data Processing

## Hospital Names

```{r,fig.width=10,fig.height=10}
balance = read_excel(paste0(input_path,'BalanceSheet.xlsx'),
                           col_types = c('numeric', 'text', rep('numeric',26)))

#name to hosid table
names_idx = seq(1,dim(balance)[1],17)
names_df = balance[names_idx,1:2]
names_df = names_df[complete.cases(names_df),]
names(names_df) = c('hosid','hosname')

datatable(names_df)
```

## Yearly Data

### Balance Sheet

```{r,fig.width=10,fig.height=10}
balance = balance[-names_idx,]
names(balance) = c('hosid','metric',names(balance)[3:dim(balance)[2]])
balance = balance[!is.na(balance$hosid),]

#melted
balance_m = melt(balance, id.vars = c('hosid','metric'))
names(balance_m) = c('hosid','metric','yrs','value')
balance_m$yrs = as.numeric(as.character(balance_m$yrs))
balance_m$metric = sapply(balance_m$metric,process_col)
print(unique(balance_m$metric))

#write
write_fst(balance_m,paste0(output_path,'balance_m.fst'))

#averge
balance_agg = balance_m %>% group_by(metric,yrs) %>% dplyr::summarise(avg_value = mean(value,na.rm = TRUE))

#sample plot
balance_sample = balance_m %>% filter(hosid==3) %>% inner_join(balance_agg)
balance_sample = melt(balance_sample,id.vars=c('hosid','metric','yrs')) %>% 
  mutate(variable=as.character(variable)) %>% filter(variable=='avg_value')
g = ggplot(balance_sample, aes(x=yrs,y=value,color=variable,group=variable)) + geom_point(size=0.5) +
  geom_line()+facet_wrap(~metric,scales='free',ncol=4) + theme_fivethirtyeight() +
  scale_color_wsj() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(1990,2016,5))
g
```

### Financial Ratios
```{r,fig.width=10,fig.height=10}
ratio = read_excel(paste0(input_path,'FinancialRatios.xlsx'),
                           col_types = c('numeric', 'text', rep('numeric',25)))

names(ratio) = c('hosid','metric',names(ratio)[3:dim(ratio)[2]])
ratio = ratio %>% filter(!is.na(hosid))

#melted
ratio_m = melt(ratio, id.vars = c('hosid','metric'))
names(ratio_m) = c('hosid','metric','yrs','value')
ratio_m$yrs = as.numeric(as.character(ratio_m$yrs))
ratio_m$metric = sapply(ratio_m$metric,process_col)
print(unique(ratio_m$metric))

#write
write_fst(ratio_m,paste0(output_path,'ratio_m.fst'))

#averge
ratio_agg = ratio_m %>% group_by(metric,yrs) %>% dplyr::summarise(avg_value = mean(value,na.rm = TRUE))

#sample plot
ratio_sample = ratio_m %>% filter(hosid==3) %>% inner_join(ratio_agg)
ratio_sample = melt(ratio_sample,id.vars=c('hosid','metric','yrs')) %>% 
  mutate(variable=as.character(variable)) %>% filter(variable=='avg_value')
g = ggplot(ratio_sample, aes(x=yrs,y=value,color=variable,group=variable)) + geom_point(size=0.5) +
  geom_line()+facet_wrap(~metric,scales='free',ncol=4) + theme_fivethirtyeight() +
  scale_color_wsj() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(1990,2016,5))
g
```

### FTE
```{r,fig.width=10,fig.height=10}
fte = read_excel(paste0(input_path,'FTE.xlsx'),
                           col_types = c('numeric', rep('numeric',41)))

names(fte) = c('hosid',names(fte)[2:dim(fte)[2]])

#melted
fte_m = melt(fte, id.vars = c('hosid'))
names(fte_m) = c('hosid','yrs','fte')
fte_m$yrs = as.numeric(as.character(fte_m$yrs))

#write
write_fst(fte_m,paste0(output_path,'fte_m.fst'))

#averge
fte_agg = fte_m %>% group_by(yrs) %>% dplyr::summarise(avg_fte = mean(fte,na.rm = TRUE))

#sample plot
fte_sample = fte_m %>% filter(hosid==3) %>% inner_join(fte_agg)
fte_sample = melt(fte_sample,id.vars=c('hosid','yrs')) %>% 
  mutate(variable=as.character(variable)) %>% filter(variable=='avg_fte')
g = ggplot(fte_sample, aes(x=yrs,y=value,color=variable,group=variable)) + geom_point(size=0.5) +
  geom_line()+ theme_fivethirtyeight() +
  scale_color_wsj() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(1975,2016,5))
g
```

### Income Statement
```{r,fig.width=10,fig.height=10}
income = read_excel(paste0(input_path,'IncomeStatement.xlsx'),
                           col_types = c('numeric', 'text', rep('numeric',26)))

names(income) = c('hosid','metric',names(income)[3:dim(income)[2]])
income = income %>% filter(!is.na(metric))

#melted
income_m = melt(income, id.vars = c('hosid','metric'))
names(income_m) = c('hosid','metric','yrs','value')
income_m$yrs = as.numeric(as.character(income_m$yrs))
income_m$metric = sapply(income_m$metric,process_col)
print(unique(income_m$metric))

#write
write_fst(income_m,paste0(output_path,'income_m.fst'))

#averge
income_agg = income_m %>% group_by(metric,yrs) %>% dplyr::summarise(avg_value = mean(value,na.rm = TRUE))

#sample plot
income_sample = income_m %>% filter(hosid==3) %>% inner_join(income_agg)
income_sample = melt(income_sample,id.vars=c('hosid','metric','yrs')) %>% 
  mutate(variable=as.character(variable)) %>% filter(variable=='avg_value')
g = ggplot(income_sample, aes(x=yrs,y=value,color=variable,group=variable)) + geom_point(size=0.5) +
  geom_line()+facet_wrap(~metric,scales='free',ncol=4) + theme_fivethirtyeight() +
  scale_color_wsj() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(1990,2016,5))
g
```

### Rate Analysis
```{r,fig.width=10,fig.height=10}
rate = read_excel(paste0(input_path,'RateAnalysis.xlsx'),
                           col_types = c('numeric', 'text', rep('numeric',25)))

names(rate) = c('hosid','metric',names(rate)[3:dim(rate)[2]])
rate = rate %>% filter(!is.na(hosid))

#melted
rate_m = melt(rate, id.vars = c('hosid','metric'))
names(rate_m) = c('hosid','metric','yrs','value')
rate_m$yrs = as.numeric(as.character(rate_m$yrs))
rate_m$metric = sapply(rate_m$metric,process_col)

#remove case mix index anomaly
remove_cmi_anomaly = function(metric,value){
  is_anomaly = value > 10 & metric=='case_mix_index'
  if (is_anomaly | is.na(is_anomaly)){
    return(NA)
  } else {
    return(value)
  }
}
rate_m$value = mapply(remove_cmi_anomaly,rate_m$metric,rate_m$value)

#write
write_fst(rate_m,paste0(output_path,'rate_m.fst'))

#averge
rate_agg = rate_m %>% group_by(metric,yrs) %>% dplyr::summarise(avg_value = mean(value,na.rm = TRUE))

#sample plot
rate_sample = rate_m %>% filter(hosid==3) %>% inner_join(rate_agg)
rate_sample = melt(rate_sample,id.vars=c('hosid','metric','yrs')) %>% 
  mutate(variable=as.character(variable)) %>% filter(variable=='avg_value')
g = ggplot(rate_sample, aes(x=yrs,y=value,color=variable,group=variable)) + geom_point(size=0.5) +
  geom_line()+facet_wrap(~metric,scales='free',ncol=4) + theme_fivethirtyeight() +
  scale_color_wsj() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(1990,2016,5))
g
```

### Volume Trends
```{r,fig.width=10,fig.height=10}
volume = read_excel(paste0(input_path,'VolumeTrends.xlsx'),
                           col_types = c('numeric', 'text', rep('numeric',25)))

names(volume) = c('hosid','metric',names(volume)[3:dim(volume)[2]])
volume = volume %>% filter(!is.na(hosid))

#get rid of names
names_idx = seq(1,dim(volume)[1],16)
volume = volume[-names_idx,]

#melted
volume_m = melt(volume, id.vars = c('hosid','metric'))
names(volume_m) = c('hosid','metric','yrs','value')
volume_m$yrs = as.numeric(as.character(volume_m$yrs))
volume_m$metric = sapply(volume_m$metric,process_col)
print(unique(volume_m$metric))

#write
write_fst(volume_m,paste0(output_path,'volume_m.fst'))

#averge
volume_agg = volume_m %>% group_by(metric,yrs) %>% dplyr::summarise(avg_value = mean(value,na.rm = TRUE))

#sample plot
volume_sample = volume_m %>% filter(hosid==3) %>% inner_join(volume_agg)
volume_sample = melt(volume_sample,id.vars=c('hosid','metric','yrs')) %>% 
  mutate(variable=as.character(variable)) %>% filter(variable=='avg_value')
g = ggplot(volume_sample, aes(x=yrs,y=value,color=variable,group=variable)) + geom_point(size=0.5) +
  geom_line()+facet_wrap(~metric,scales='free',ncol=4) + theme_fivethirtyeight() +
  scale_color_wsj() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(1990,2016,5))
g
```


### Putting It Together

Out of all datasets, we construct a consolidated dataset based on selected features. We also add denominator defined as follows:

$$denominator = caseMixIndex * admissions$$

```{r}
#selected columns
income_selected = c('inpatient_revenue','outpatient_revenue','total_patient_revenue',
                    'bad_debts','contractual_adjustments',
                    'charity_care','other_adjustments','other_operating_revenue',
                    'tax_revenue','salaries_and_benefits','employee_benefits','professional_fees',
                    'supplies','purchased_services_utilities','purchased_services_other','depreciation',
                    'rentals_and_leases','insurance','license_and_taxes','interest','other_direct_expenses',
                    'non_operating_income','extraordinary_items','federal_income_tax',
                    #for y
                    'net_revenue_or_expense')

balance_selected = c('total_current_assets','total_board_designated_assets','total_property_plant_equipment',
                    'less_accumulated_depreciation','total_investments_other_assets','total_intangible_assets',
                    'total_current_liabilities','total_deferred_credits','total_long_term_debt',
                    'less_current_maturities_of_ltd','unrestricted_fund_balance')
     
volume_selected = c('admissions','births','patient_days','lic_beds',
                    'avail_beds','intensive_care_patient_days','acute_care_patient_days',
                    'surgical_services_op_min','laboratory_cap_units','radiology_rvus',
                    'ct_scanning_hect_units','emergency_room_visits')

rate_selected = c('case_mix_index')

ratio_selected = c('deductible','markup','days_in_patient_ar','average_payment_period','average_age_of_plant',
                  #for y
                  'operating_margin','return_on_total_assets')

#function to generate features
# selected_col = income_selected
# fst_name = paste0(output_path,'income_m.fst')
feature_spread = function(fst_name){
    #load
  df_m = read_fst(fst_name) %>% 
    mutate(hosid_yrs = paste(hosid,yrs,sep='_')) %>%
    select(hosid_yrs,metric,value)
  
  df_s = df_m %>% spread(metric,value) %>%
    mutate(yrs = sapply(hosid_yrs, FUN=function(x) strsplit(x,'_')[[1]][2]),
           hosid = sapply(hosid_yrs, FUN=function(x) strsplit(x,'_')[[1]][1])) %>% 
    ungroup %>% select(-hosid_yrs) 
  return(df_s)
}

feature_denom = function(df_s,denom_s){
  
  joined = denom_s %>% inner_join(df_s)
  
  for(i in 4:dim(joined)[2]){
    joined[,i] = mapply(function(x,y) x / y, joined[,i], joined[,3])
  }
  
  joined = joined %>% select(-denominator)
  return(joined)
}

feature_lag = function(df_s,selected_col,lag_list=c(1,2,3)){
  
  df_s = df_s[,c('yrs','hosid',selected_col)] %>% data.table
  for (l in lag_list){
    for (metric in selected_col){
    df_s = df_s[,paste0(metric,'_',l):=lag(eval(parse(text=metric)),l), by=hosid]
    }
  }
  
  return(df_s)
}

feature_growth = function(df_s,selected_col, lag_list=c(1,2,3)){
  #growth
  df_s = df_s %>% data.frame
  for (l in lag_list){
    for (metric in selected_col){
      df_s[,paste0(metric,'_g',l)] = mapply(gen_growth,df_s[,metric],df_s[,paste0(metric,'_',l)])
    } 
  }
  
  return(df_s)
}
```


```{r}
#create denominator
denom_s = feature_spread(paste0(output_path,'rate_m.fst'))
denom_s$denominator = denom_s$case_mix_index * denom_s$admissions
denom_s = denom_s %>% select(hosid,yrs,denominator)

#create those with denom
#income
income_s = feature_spread(paste0(output_path,'income_m.fst')) %>% mutate(hosid = as.character(hosid), yrs = as.character(yrs)) 
#income_s = feature_denom(income_s,denom_s)
# income_s = feature_lag(income_s,income_selected)
# income_s = feature_growth(income_s,income_selected)
#balance
balance_s = feature_spread(paste0(output_path,'balance_m.fst')) %>% mutate(hosid = as.character(hosid), yrs = as.character(yrs)) 
#balance_s = feature_denom(balance_s,denom_s)
# balance_s = feature_lag(balance_s,balance_selected)
# balance_s = feature_growth(balance_s,balance_selected)
#volume
volume_s = feature_spread(paste0(output_path,'volume_m.fst')) %>% mutate(hosid = as.character(hosid), yrs = as.character(yrs)) 
#volume_s = feature_denom(volume_s,denom_s)
# volume_s = feature_lag(volume_s,volume_selected)
# volume_s = feature_growth(volume_s,volume_selected)

#create those without denom
#rate
rate_s = feature_spread(paste0(output_path,'rate_m.fst')) %>% mutate(hosid = as.character(hosid), yrs = as.character(yrs)) 
# rate_s = feature_lag(rate_s,rate_selected)
# rate_s = feature_growth(rate_s,rate_selected)
#ratio
ratio_s = feature_spread(paste0(output_path,'ratio_m.fst')) %>% mutate(hosid = as.character(hosid), yrs = as.character(yrs)) 
# ratio_s = feature_lag(ratio_s,ratio_selected)
# ratio_s = feature_growth(ratio_s,ratio_selected)

#already spread
fte_s = read_fst(paste0(output_path,'fte_m.fst'))
fte_s = fte_s %>% mutate(hosid = as.character(hosid), yrs = as.character(yrs)) 
#fte_s = feature_denom(fte_s,denom_s)
# fte_s = feature_lag(fte_s,'fte')
# fte_s = feature_growth(fte_s,'fte')

# denom_s = feature_lag(denom_s,'denominator')
# denom_s = feature_growth(denom_s,'denominator')

df = denom_s %>% inner_join(income_s, by=c('yrs','hosid')) %>% inner_join(balance_s, by=c("yrs", "hosid")) %>% 
  inner_join(volume_s, by=c("yrs", "hosid")) %>%
  inner_join(rate_s %>% 
             select(-tax_revenue,-total_patient_revenue,-admissions,-patient_days), 
             by=c("yrs", "hosid")) %>% 
  inner_join(ratio_s, by=c("yrs", "hosid")) %>% inner_join(fte_s, by=c("yrs", "hosid"))
df = df[complete.cases(df),] 
#filter for those that has at least 20 data points (15 train and 5 test)
enough_data = df %>% group_by(hosid) %>% summarise(nb_yrs = n()) %>% filter(nb_yrs>=20) %>%
  select(hosid)
df = df %>% inner_join(enough_data,by=c('hosid')) %>% data.table
df = df[,yrs_id:= row_number(desc(yrs)),by=hosid] %>% mutate(test_flag = ifelse(yrs_id <= 5,1,0))
```

## Quarterly Data

```{r}
quarterly = read_excel(paste0(input_path,'quarterly.xlsx'),
                           col_types = c('numeric', 'numeric','text', rep('numeric',107)))
names(quarterly) = c('yrs','hosid','hosname','quarter','licensed_beds',
#admits
'acute_care_medicare_admits','acute_care_medicaid_admits','acute_care_other_admits','acute_care_total_admits',
'swing_bed_medicare_admits','swing_bed_medicaid_admits','swing_bed_other_admits','swing_bed_total_admits',
'snf_medicare_admits','snf_medicaid_admits','snf_other_admits','snf_total_admits',
'psych_medicare_admits','psych_medicaid_admits','psych_other_admits','psych_total_admits',
#patient days
'acute_care_medicare_patient_days','acute_care_medicaid_patient_days','acute_care_other_patient_days','acute_care_total_patient_days',
'swing_bed_medicare_patient_days','swing_bed_medicaid_patient_days','swing_bed_other_patient_days','swing_bed_total_patient_days',
'snf_medicare_patient_days','snf_medicaid_patient_d ays','snf_other_patient_days','snf_total_patient_days',
'psych_medicare_patient_days','psych_medicaid_patient_days','psych_other_patient_days','psych_total_patient_days',
#surgeries and births
'total_inpatient_surgeries','total_births','total_newborn_days','emergency_room_admits',
#visits; all sums to outpatient visits
'emergency_room_visits','outpatient_surgery_visits', 'observation_visits','home_health_visits',
'all_other_visits','total_outpatient_visits',
#patient revenue
'acute_medicare_inpatient_revenue','acute_medicaid_inpatient_revenue','acute_other_inpatient_revenue',
'acute_inpatient_total_revenue',
'acute_medicare_outpatient_revenue','acute_medicaid_outpatient_revenue',
'acute_other_outpatient_revenue','acute_outpatient_total_revenue',
'swing_bed_medicare_revenue','swing_bed_medicaid_revenue','swing_bed_other_revenue','swing_bed_total_revenue',
'snf_medicare_revenue','snf_medicaid_revenue','snf_other_revenue','snf_total_revenue',
'psych_medicare_revenue','psych_medicaid_revenue','psych_other_revenue','psych_total_revenue',
'home_health_medicare_revenue','home_health_medicaid_revenue','home_health_other_revenue','home_health_total_revenue',
'medicare_patient_revenue','medicaid_patient_revenue','other_patient_revenue','total_patient_revenue',
#contractuals
'acute_contractuals_medicare','acute_contractual_medicaid','acute_contractual_other','acute_contractual_total',
'other_contractuals_medicare','other_contractual_medicaid','other_contractual_other','other_contractual_total',
'total_contractuals_medicare','total_contractual_medicaid','total_contractual_other','total_contractual_total',
#charity
'total_charity_care',
#employees
'facility_payroll_expenses','facility_hours','physician_payroll_expenses','physician_hours',
'total_payroll_expenses',
'total_hours','employee_benefit_expenses',
#expenses
'supply_expenses','depre_expenses','interest_expenses','bad_debts_expenses','other_operating_expenses',
'total_operating_expenses',
#other reveneue
'other_operating_revenue',
#margin
'operating_margin',
#non operating
'net_non_operating',
#tax
'tax_revenue',
#net margin
'net_margin',
#account receivables
'gross_account_receivables_medicare','gross_account_receivables_medicaid',
'gross_account_receivables_other','gross_account_receivables_total'
)
#ds
quarterly$ds = paste0(quarterly$yrs,'-',sprintf('%02s',quarterly$quarter*3),'-01')
#test_flag
#filter for those that has at least 88 data points 
enough_data = quarterly %>% group_by(hosid) %>% summarise(nb_yrs = n()) %>% filter(nb_yrs>=88) %>%
  select(hosid)
quarterly = quarterly %>% inner_join(enough_data,by=c('hosid')) %>% data.table
#get 20 train samples and 20 test samples
quarterly = quarterly[,id:= row_number(desc(ds)),by=hosid] %>% mutate(test_flag = ifelse(id <= 20,1,0)) %>%
  mutate(train_flag = ifelse(id > 20 & id <= 40,1,0))

#filter out medicare/medicaid
quarterly = quarterly[,!grepl('medicare|medicaid',names(quarterly))]

quarterly = quarterly %>% 
  #redefine revenue
  mutate(
  #chronic inpatient = swing + snf + psych
  #patient revenue = outpatient&home + chronic_inpatient + acute_inpatient
  #total revenue = patient revenue + other operating revenue
  outpatient_home_health_total_revenue = acute_outpatient_total_revenue + home_health_total_revenue,
  chronic_total_admits = psych_total_admits + swing_bed_total_admits + snf_total_admits,
  chronic_inpatient_total_revenue = psych_total_revenue + swing_bed_total_revenue + snf_total_revenue,
  total_revenue = total_patient_revenue + other_operating_revenue) %>%
  #efficiency metrics
  mutate(
  #revenue per times
  outpatient_home_health_per = outpatient_home_health_total_revenue / total_outpatient_visits,
  acute_inpatient_per = acute_inpatient_total_revenue / acute_care_total_admits,
  chronic_inpatient_per = chronic_inpatient_total_revenue / chronic_total_admits,
  #patient expenses per revenue
  contractual_eff = total_contractual_total / total_patient_revenue,
  physician_eff = physician_payroll_expenses / total_patient_revenue,
  facility_eff = facility_payroll_expenses / total_patient_revenue,
  benefits_eff = employee_benefit_expenses / total_patient_revenue,
  bad_debts_eff = bad_debts_expenses / total_patient_revenue,
  supply_eff = supply_expenses / total_patient_revenue,
  #other expenses per revenue
  interest_eff = interest_expenses / total_patient_revenue,
  depre_eff = depre_expenses / total_patient_revenue,
  other_eff = other_operating_expenses / total_patient_revenue,
  charity_eff = total_charity_care / total_patient_revenue
  ) %>% 
  #pure margin
  mutate(
  #operating expense = payroll + benefits + bad debts + supply + (others + depre + interest)
  #patient expense = payroll + benefits + bad debts + supply + (contractual)
  total_patient_expenses = total_payroll_expenses+ employee_benefit_expenses + bad_debts_expenses +
    supply_expenses + total_contractual_total,
  #operating margin = total revenue - (contractual + charity + operating expense)
  #patient margin = patient revenue - patient expense
  patient_margin=total_patient_revenue - total_patient_expenses)

#get total asset
yearly = df %>% select(hosid,yrs,total_assets) %>%
  mutate_all(funs(as.numeric))
quarterly = quarterly %>% left_join(yearly) %>%
  mutate(roa = net_margin / total_assets)

write_fst(quarterly,paste0(output_path,'quarterly.fst'))
```

# Models

```{r}
#train and test
train_df = quarterly  %>% filter(train_flag == 1)
test_df = quarterly %>% filter(test_flag == 1)
show_cols = c('patient_margin','total_patient_revenue','total_patient_expenses',
              'outpatient_home_health_total_revenue',
              'total_outpatient_visits','acute_inpatient_total_revenue',
              'acute_care_total_admits','chronic_inpatient_total_revenue',
              'chronic_total_admits','total_payroll_expenses','total_hours',
              'employee_benefit_expenses','bad_debts_expenses','supply_expenses',
              'total_contractual_total'
              )
mape_overalls = NULL 
mape_groups = NULL


for(col in show_cols){
  #set up dependent variable
  y = col
  xs = c()
  results = NULL

  for(the_id in enough_data$hosid){
    #create df for each hospital
    sample_train = train_df %>% filter(hosid==the_id) 
    sample_test = test_df %>% filter(hosid==the_id)
    
    #create prophet dfs
    prophet_train = data.frame(ds = sample_train$ds, 
                               y = sample_train[,y])
    prophet_test = data.frame(ds = sample_train$ds,
                              y_actual = sample_test[,y])
    
    #initialize model
    m = prophet(n.changepoints = 5)
    
    #add regressors
    for (x in xs){
      prophet_train[,x] = sapply(sample_train[,x],function(x) x + rnorm(1))
      prophet_test[,x] = sapply(sample_test[,x],function(x) x + rnorm(1))
      m = m %>% add_regressor(x)
    }
    
    #fit
    m = fit.prophet(m, prophet_train)
    
    #create data for predictions
    future = make_future_dataframe(m, periods = 20, freq = 'quarter') 
    
    #add regressors
    for (x in xs){
      future[,x] = c(prophet_train[,x],prophet_test[,x])
    }
    #get results
    forecast = predict(m, future)
    #plot(m,forecast)
    forecast$ds = as.character(forecast$ds)
    header_df = data.frame(hosid = sample_test$hosid,
                           ds =sample_test$ds,id = sample_test$id, 
                           y_actual = prophet_test$y_actual)
    result = header_df %>% inner_join(forecast)
    result$APE = mapply(APE,result$y_actual,result$yhat)
    results = rbind(results,result)
  }
  
  mape_group = results %>% group_by(id) %>% summarise(nb_pred = n(),MAPE= mean(APE,na.rm = TRUE), 
                                                          MdAPE = median(APE,na.rm = TRUE)) %>%
    ungroup %>% mutate(variable=y)
  mape_overall = results %>% summarise(nb_pred = n(),MAPE= mean(APE,na.rm = TRUE), 
                                       MdAPE = median(APE,na.rm = TRUE)) %>% ungroup %>%
    mutate(variable=y)
  mape_overalls = rbind(mape_overalls,mape_overall)
  mape_groups = rbind(mape_groups,mape_group)
}

write_fst(mape_overalls,paste0(output_path,'mape_overalls.fst'))
#mape_overalls = read_fst(paste0(output_path,'mape_overalls.fst'))
```


