---
title: "Continuum Enterprise"
resource_files:
- data/processed/quarterly.fst
- data/processed/mape_overalls.fst
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: row
---


```{r setup, include=FALSE}
#comments
#comparison between 10M and 5M OPD buildings; MPV IRR q

les_packages <- lapply(c('prophet','data.table','tidyverse','ggplot','ggthemes','DT',
                         'reshape2','fst','plotly','DT','ggthemes'),require, character.only = TRUE)
## load files
output_path =  '../data/processed/'

APE = function(actual,pred){
  if(is.nan(actual) | is.na(actual)) return(NA)
  if(actual!=0){
    return(abs(pred-actual) / abs(actual))
  } else {
    return(NA)
  }
}


gen_forecast = function(df,y,periods=20, l1= 0.05){
  #train_test
  sample_train = df %>% filter(train_flag==1) 
  sample_test = df %>% filter(test_flag==1)
  #create prophet dfs
  prophet_train = data.frame(ds = sample_train$ds, y = sample_train[,y])
  prophet_test = data.frame(ds = sample_train$ds,y_actual = sample_test[,y])
      
  #initialize model
  m = prophet(changepoint.prior.scale = l1)
      
  #fit
  m = fit.prophet(m, prophet_train)
      
  #create data for predictions
  future = make_future_dataframe(m, periods = periods, freq = 'quarter') 
  
  #get results
  forecast = predict(m, future)
  #plot(m,forecast)
  forecast$ds = as.character(forecast$ds)
  header_df = data.frame(hosid = df$hosid,
                         ds = df$ds,
                         id = df$id,
                         y_actual = df[,y],
                         test_flag = ifelse(df$test_flag==1,TRUE,FALSE))
  result = header_df %>% inner_join(forecast) %>%
    mutate(APE = mapply(APE,y_actual,yhat))
  
  mape_group = result %>% filter(test_flag==1) %>% select(ds,y_actual,yhat,APE)
  
  mape_overall = result %>% filter(test_flag==1) %>% summarise(nb_pred = n(),
                                      MAPE= mean(APE,na.rm = TRUE),
                                      MdAPE = median(APE,na.rm = TRUE))
  
  return(list(result,mape_group,mape_overall))
}

plot_prophet = function(result,w=640,h=290,scaler=1/1e6){
  g = ggplot(result,aes(x=ds,y=y_actual,color=test_flag)) + geom_point() +
    geom_line(data=result,aes(x=ds,y=yhat,group=1)) + 
    geom_ribbon(data=result,aes(ymin=yhat_lower,ymax=yhat_upper,group=1),alpha=0.3) +
    theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x*scaler) +
    scale_x_discrete(breaks=result$ds[seq(1,length(result$ds),4)],
                     labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = 'none') 
  ggplotly(g,width=w,height=h)
}

#load dataframe
quarterly = read_fst(paste0(output_path,'quarterly.fst'))

#perfect fits
#0.05 - 29, 3
#fix
#0.05 - 23
examples = data.frame(hosid = c(3,29,23)) %>% inner_join(quarterly %>% group_by(hosid) %>% summarise(hosname=min(hosname)) %>% ungroup)
```

Sidebar {.sidebar}
=====================================

**ADJUST TARGETS:**

```{r}
sliderInput('total_patient_revenue', 'Patient Revenues (%)', -100,100,0,10)

sliderInput('total_payroll_expenses', 'Payroll (%)', -100,100,0,10)

sliderInput('employee_benefit_expenses', 'Employee Benefits (%)',-100,100,0,10)

sliderInput('supply_expenses', 'Supplies and Utilities (%)', -100,100,0,10)

sliderInput('total_contractual_total', 
            'Contractual Adjustments (%)', -100,100,0,10)
```


Overall
=====================================

Row 
-----------------------------------------------------------------------

### Scenario

```{r}
selectInput('scenario', 'Pick Hospital', examples$hosname)
scenario_y = c('patient_margin','total_patient_revenue','total_patient_expenses')

df = reactive({
  quarterly %>% 
    filter(hosid==examples[examples$hosname==input$scenario,'hosid'] & (train_flag==1 | test_flag ==1))
})
```

Forecast Statistics

```{r}
renderDataTable({
  mape_overalls = NULL
  for (y in scenario_y){
    mape_overall = gen_forecast(df(),y)[[3]] %>% mutate(variable = y) %>%
      select(variable,nb_pred,MAPE,MdAPE)
    mape_overalls = rbind(mape_overalls,mape_overall)
  }
  datatable(mape_overalls,
                    colnames = c('Variable' = 1, 
                       '# Forecasts' = 2 ,
                       'Mean Error' = 3 ,
                       'Median Error' = 4 
                       ),
          class = 'cell-border stripe' , 
          rownames = FALSE,
          options=list(dom='t')
          ) %>% 
  formatPercentage(c('Mean Error','Median Error'), 2)
})
```

### Patient Margin

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'patient_margin')[[1]])
})
```

Row 
-----------------------------------------------------------------------

### Total Patient Revenue

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_patient_revenue')[[1]])
})
```

### Total Patient Expenses

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_patient_expenses')[[1]])
})
```


Revenue
=====================================


Row 
-----------------------------------------------------------------------

### Inpatient Revenue

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'inpatient_total_revenue')[[1]],w=420,h=175)
})
```

### Inpatient Admits

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'inpatient_total_admits')[[1]],w=420,h=175)
})
```

### Inpatient Revenue per Admit

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'inpatient_per')[[1]],w=420,h=175)
})
```

Row 
-----------------------------------------------------------------------

### Outpatient Revenue

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'outpatient_total_revenue')[[1]],w=420,h=175)
})
```

### Outpatient Visits

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_outpatient_visits')[[1]],w=420,h=175)
})
```

### Outpatient Revenue per Visit

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'outpatient_per')[[1]],w=420,h=175)
})
```

Row 
-----------------------------------------------------------------------

### Total Patient Days

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_patient_days')[[1]],w=420,h=175)
})
```

### Average Patient Days

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_patient_days')[[1]],w=420,h=175)
})
```

### Revenue per Patient Day

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'revenue_per_patient_days')[[1]],w=420,h=175)
})
```



Expenses
=====================================

Row 
-----------------------------------------------------------------------

### **Payroll Expenses**

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_payroll_expenses')[[1]],w=420,h=175)
})
```

### **Employee Benefits Expenses**

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'employee_benefit_expenses')[[1]],w=420,h=175)
})
```

### Bad Debts Expenses

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'bad_debts_expenses')[[1]],w=420,h=175)
})
```

Row 
-----------------------------------------------------------------------

### **Supply Expenses**

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'supply_expenses')[[1]],w=420,h=175)
})
```

###  **Contractual Adjustments**

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_contractual_total')[[1]],w=420,h=175)
})
```

### Depreciation Expenses

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'depre_expenses')[[1]],w=420,h=175)
})
```

Row 
-----------------------------------------------------------------------

### Interest Expenses

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'interest_expenses')[[1]],w=420,h=175)
})
```

### Charity Care

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'total_charity_care')[[1]],w=420,h=175)
})
```

### Other Operating Expenses

```{r}
renderPlotly({
  plot_prophet(gen_forecast(df(),'other_operating_expenses')[[1]],w=420,h=175)
})
```


Scorecard
=====================================

```{r}
patient_expenses = c('total_payroll_expenses','employee_benefit_expenses',
            'supply_expenses','total_contractual_total')

margin_call = reactive({
  control_df = df()[,c('ds','patient_margin','total_patient_expenses',
                       'total_operating_expenses',
                       'operating_margin','net_margin','total_assets',
                       'total_patient_revenue','total_revenue')]
  
  #patient expenses
  alt_df = df()[,c('ds','test_flag')]
  for(i in patient_expenses){
    input_i = eval(parse(text=paste0('input$',i)))
    alt_df[,i] = gen_forecast(df(),i)[[1]][,'yhat'] * (1 + input_i/100)
  }
  
  #other expenses
  alt_df$bad_debts_expenses = df()$bad_debts_expenses
  alt_df$interest_expenses = df()$interest_expenses + 
    input$add_interest/100/4 * input$add_ppe * 1e6
  alt_df$depre_expenses = df()$depre_expenses + input$add_ppe * 1e6 / 20
  alt_df$total_charity_care = df()$total_charity_care
  alt_df$other_operating_expenses = df()$other_operating_expenses + 
    input$add_other_operating * 1e6 / 4
  alt_df = alt_df %>% mutate(total_operating_expenses = 
                               total_payroll_expenses+employee_benefit_expenses+
                               supply_expenses+bad_debts_expenses+other_operating_expenses+
                               depre_expenses+interest_expenses,
                            total_patient_expenses=total_payroll_expenses+ 
                              employee_benefit_expenses + 
                              supply_expenses + total_contractual_total)
  #asset
  alt_df$total_assets = df()$total_assets + input$add_ppe * 1e6
  #total revenue
  alt_df$total_patient_revenue = 
    gen_forecast(df(),'total_patient_revenue')[[1]][,'yhat'] * 
    (1 + input$total_patient_revenue/100)
  alt_df$total_revenue = alt_df$total_patient_revenue +
    df()$other_operating_revenue
  alt_df$net_non_operating = df()$net_non_operating
  alt_df$tax_revenue = df()$tax_revenue
  #calculate metrics
  alt_df = alt_df %>%
    mutate(patient_margin = total_patient_revenue - total_patient_expenses,
           operating_margin = total_revenue - (total_operating_expenses + 
                                               total_contractual_total +
                                               total_charity_care),
           net_margin = operating_margin + net_non_operating + tax_revenue)
  alt_df = alt_df[,
    c('ds','test_flag','patient_margin','operating_margin','total_operating_expenses',
      'total_patient_expenses','total_patient_revenue','total_revenue'
      )]
  
  #plot df
  alt_m = melt(alt_df,id.vars=c('ds','test_flag')) %>% 
    filter(test_flag==1) %>% select(-test_flag) %>%
    mutate(control_flag='alternative')
  
  control_m = melt(control_df,id.vars=c('ds')) %>% mutate(control_flag = 'control')
  
  margin_df = rbind(control_m,alt_m)
  margin_df
})
```

Row
-------------------------------------

### Patient Revenue

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='total_patient_revenue')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=420,height=290)
})
```

### Patient Expenses

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='total_patient_expenses')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=420,height=290)
})
```

### Patient Margin

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='patient_margin')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=420,height=290)
})
```

Row
-------------------------------------

### Total Operating Revenue

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='total_revenue')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=420,height=290)
})
```

### Total Operating Expenses

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='total_operating_expenses')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=420,height=290)
})
```

### Operating Margin

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='operating_margin')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=420,height=290)
})
```


Expansion
=====================================

Row
-------------------------------------

### Investment

```{r}
fluidRow(
      column(6,
        numericInput('add_ppe','Add PPE (M)',value = 0),
        numericInput('add_interest','Interest Rate (%)',value = 7),
        numericInput('add_other_operating','Other Additional Expenses (M)',
                     value = 0)
      ),
      column(6,
        numericInput('discount_rate','Discount Rate (%)',value = 12),
        renderText({
          add_revenues = tail(gen_forecast(df(),
            'total_patient_revenue')[[1]][,'yhat'] * input$total_patient_revenue/100,20)
          sum_revenue = 0
          for(i in 1:length(add_revenues)){
            add_cashflow = add_revenues[i] - input$add_other_operating * 1e6 / 4 -
              input$add_interest/100/4 * input$add_ppe * 1e6
            sum_revenue = sum_revenue + (add_cashflow/(1+input$discount_rate)^(i/4))
          }
          paste0('5-year Net Present Value: ', 
                 round((sum_revenue - input$add_ppe * 1e6)/1e6,2),'M')
        })
      )
)
```

### Operating Margin

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='operating_margin')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=680,height=290)
})
```

Row
-------------------------------------

### Operating Revenue

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='total_revenue')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=680,height=290)
})
```

### Operating Expenses

```{r}
renderPlotly({
  mar = margin_call() %>% filter(variable=='total_operating_expenses')
  g = ggplot(mar,aes(x=ds,y=value,color=control_flag, group=control_flag)) + 
    geom_point() + geom_line() + theme_fivethirtyeight() + scale_color_wsj() +
    scale_y_continuous(labels=function(x) x/1e6) +
    scale_x_discrete(breaks=mar$ds[seq(1,length(mar$ds),4)],labels=function(x) substr(x,1,4)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.title=element_blank())
  ggplotly(g,width=680,height=290)
})
```


Model Performance
=====================================

Row
-------------------------------------

### Mean and Median Absolute Percentage Errors

```{r}
mape_performance = read_fst(paste0(output_path,'mape_overalls.fst')) %>%
  select(variable,nb_pred,MAPE,MdAPE)
datatable(mape_performance %>% arrange(MdAPE),
                  colnames = c('Variable' = 1, 
                     '# Forecasts' = 2 ,
                     'Mean Error' = 3 ,
                     'Median Error' = 4 
                     ),
        class = 'cell-border stripe' , 
        rownames = FALSE, options=list(pageLength = 15)
        ) %>% 
formatPercentage(c('Mean Error','Median Error'), 2)
```